Assessing \Htwo{} and \COtwo{} Network Strategies in Europe

### 2023-12-15 report/submissions/journal/paper/paper.tex
paper: remove footnote, add more figures to appendix

In summary, both network models prefer CC from point sources distributed across the continent as well as electrolysis concentrated in regions with abundant renewable resources. The two network approaches reveal fundamentally different strategies to integrate these low-cost resources into the system: The \carbongridmodel{} transports and sequesters low-cost carbon from distributed point sources and co-locates CU with low-cost electrolysis and DAC. The \hydrogengridmodel{} distributes CU alongside with carbon point sources to which low-cost hydrogen from regions with abundant renewable resources is transported.
%
% However, differences between the two models are notable for the placement of CU, CS and DAC due to their distinct \COtwo{} and \Htwo{} price patterns. CU locations are more distributed in the \modH{}, while they cluster in a few valleys of low-cost \Htwo{} in the \modCO{}. CS co-locates in the regions with the lowest \COtwo{} costs (among those with offshore sequestration potential), but these regions are very different between the models. DAC is placed in regions with the best renewable resources in both models, but it is used for only CU in the \modCO{} and for both CU and CS in the \modH{}.
Finally, the system cost advantage of the \modH{} stems from being able to supply the exogenous hydrogen demands in Central Europe with low-cost hydrogen imports, while the \modCO{} is forced to deploy electrolysis in regions with poor renewables resources as well as additional gas infrastructure in Central Europe.

@@

To supply carbon needed for CS and CU, the system can choose to deploy carbon capture (CC) technologies at various point sources of fossil or biogenic origin (see below), or through DAC facilities.
The concept of a merit order for captured carbon plays a pivotal role in optimizing the deployment of carbon capture technologies based on their economic feasibility.
This merit order ranks all CC technologies according to their relative costs to capture an additional marginal ton\footnote[2]{The cost value for a technology depends on the exact system configuration and may vary by scenario and even by region, since many technologies produce other (primary) outputs such as electricity and heat. The costs to capture an additional marginal ton of carbon may also be different from the average cost of capture.}
At the lower end of the cost spectrum, CC technologies applied to process emissions, such as those from cement, offer a cost-effective starting point.
Following this, biomass combined heat and power (CHP) systems provide a dual benefit of energy production and carbon capture.
Moving up the scale, gas used in industrial applications and biomass employed for industrial processes represent more costly yet viable options to capture carbon.
DAC, an emerging technology capable of extracting \COtwo{} directly from the atmosphere, stands at a higher cost level due to its current high capital costs and energy demand.
Finally, based on our model results, biogas upgrading, a process that refines biogas to natural gas quality, incurs the highest costs in the merit order per marginal ton of captured \COtwo{}.
Biogas input is a high-cost fuel which based on the endogenous modeling decisions is not used in large quantities.
To the extent that biogas-to-gas facilities are used by the model to supply additional (carbon-neutral) gas as fuel, adding CC infrastructure incurs low costs (and thus ranges on the left-hand side of the merit order curve).
However, the price of capturing additional \COtwo{} from biogas-to-gas upgrading is high because a substantial amount of the biogas fuel costs factors into the marginal cost of captured \COtwo{}.
This merit order framework is crucial for strategically deploying necessary carbon capture solutions while balancing economic considerations.
If we consider the spatial aspect of distributed carbon capture potentials, the \COtwo{} network can exploit comparative cost advantages between CO2 bidding zones to utilize the lowest-cost CC facilities across the continent.
We assume a capture rate of 90\% for CC on process emissions, SMR, biogas-to-gas, as well as gas and biomass used in industry, and 95\% for CC on biomass and gas CHPs.

@@

\begin{figure}[ht]
    \centering
    \includegraphics[width=\linewidth]{comparison/emission-reduction-full/figures/90_nodes/cost_bar_transmission.png}
    \caption{Annual transmission system cost as a function of the net carbon removal scenarios considered in the study.}
    \label{fig:cost_bar_transmission}
\end{figure}

@@

>>>>

In summary, both network models prefer CC from point sources distributed across the continent as well as electrolysis concentrated in regions with abundant renewable resources. The two network approaches reveal fundamentally different strategies to integrate these low-cost resources into the system: The \carbonmodel{} transports and sequesters low-cost carbon from distributed point sources and co-locates CU with low-cost electrolysis and DAC. The \hydrogenmodel{} distributes CU alongside with carbon point sources to which low-cost hydrogen from regions with abundant renewable resources is transported.
%
% However, differences between the two models are notable for the placement of CU, CS and DAC due to their distinct \COtwo{} and \Htwo{} price patterns. CU locations are more distributed in the \modH{}, while they cluster in a few valleys of low-cost \Htwo{} in the \modCO{}. CS co-locates in the regions with the lowest \COtwo{} costs (among those with offshore sequestration potential), but these regions are very different between the models. DAC is placed in regions with the best renewable resources in both models, but it is used for only CU in the \modCO{} and for both CU and CS in the \modH{}.
Finally, the system cost advantage of the \modH{} stems from being able to supply the exogenous hydrogen demands in Central Europe with low-cost hydrogen imports, while the \modCO{} is forced to deploy electrolysis in regions with poor renewables resources as well as additional gas infrastructure in Central Europe.

@@

To supply carbon needed for CS and CU, the system can choose to deploy carbon capture (CC) technologies at various point sources of fossil or biogenic origin (see below), or through DAC facilities.
The concept of a merit order for captured carbon plays a pivotal role in optimizing the deployment of carbon capture technologies based on their economic feasibility.
This merit order ranks all CC technologies according to their relative costs to capture an additional marginal ton. The cost value for a technology depends on the exact system configuration and may vary by scenario and even by region, since many technologies produce other (primary) outputs such as electricity and heat. The costs to capture an additional marginal ton of carbon may also be different from the average cost of capture.
At the lower end of the cost spectrum, CC technologies applied to process emissions, such as those from cement, offer a cost-effective starting point.
Following this, biomass combined heat and power (CHP) systems provide a dual benefit of energy production and carbon capture.
Moving up the scale, gas used in industrial applications and biomass employed for industrial processes represent more costly yet viable options to capture carbon.
DAC, an emerging technology capable of extracting \COtwo{} directly from the atmosphere, stands at a higher cost level due to its current high capital costs and energy demand.
Finally, based on our model results, biogas upgrading, a process that refines biogas to natural gas quality, incurs the highest costs in the merit order per marginal ton of captured \COtwo{}.
Biogas input is a high-cost fuel which based on the endogenous modeling decisions is not used in large quantities.
To the extent that biogas-to-gas facilities are used by the model to supply additional (carbon-neutral) gas as fuel, adding CC infrastructure incurs low costs (and thus ranges on the left-hand side of the merit order curve).
However, the price of capturing additional \COtwo{} from biogas-to-gas upgrading is high because a substantial amount of the biogas fuel costs factors into the marginal cost of captured \COtwo{}.
This merit order framework is crucial for strategically deploying necessary carbon capture solutions while balancing economic considerations.
If we consider the spatial aspect of distributed carbon capture potentials, the \COtwo{} network can exploit comparative cost advantages between CO2 bidding zones to utilize the lowest-cost CC facilities across the continent.
We assume a capture rate of 90\% for CC on process emissions, SMR, biogas-to-gas, as well as gas and biomass used in industry, and 95\% for CC on biomass and gas CHPs.

@@

\begin{figure}[ht]
    \centering
    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/cost_bar_transmission.png}
    \caption{Annual transmission system cost for the different models in the net-zero emission scenario.}
    \label{fig:cost_bar_transmission}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics[width=\linewidth]{comparison/net-negative-0.1/figures/90_nodes/cost_bar_transmission.png}
    \caption{Annual transmission system cost for the different models in the net-negative emissions scenario.}
    \label{fig:cost_bar_transmission_nn}
\end{figure}

@@

\begin{figure*}[ht!]
    \centering
    \begin{subfigure}{.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{net-negative-0.1/co2-only/figures/90_nodes/balance_map_carbon.png}
        \label{fig:capacity_map_carbon_co2_nn}
    \end{subfigure}%
    \begin{subfigure}{.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{net-negative-0.1/co2-only/figures/90_nodes/balance_map_hydrogen.png}
        \label{fig:capacity_map_hydrogen_co2_nn}
    \end{subfigure}
    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors for the \carbonmodel{} in the net negative scenario. For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices. Carbon Sequestration offshore are drawn as full circles. Lines and arrows show the interregional transportation. The \carbongrid{} reveals a fundamentally different layout than in the Net-Zero scenario.
    }
    \label{fig:capacity_maps_co2_nn}
\end{figure*}

\begin{figure*}[ht!]
    \centering
    \begin{subfigure}{.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{net-negative-0.1/h2-only/figures/90_nodes/balance_map_carbon.png}
        \label{fig:capacity_map_carbon_h2_nn}
    \end{subfigure}%
    \begin{subfigure}{.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{net-negative-0.1/h2-only/figures/90_nodes/balance_map_hydrogen.png}
        \label{fig:capacity_map_hydrogen_h2_nn}
    \end{subfigure}
    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors for the \hydrogenmodel{} in the net negative scenario. For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices. Carbon Sequestration offshore are drawn as full circles. Lines and arrows show the interregional transportation. With the tightened emission target, the \hydrogengrid{} expands the \hydrogengrid{} layout of the Net-Zero scenario model results while increasing DAC facilities at the coast.
    }
    \label{fig:capacity_maps_h2_nn}
\end{figure*}

\begin{figure*}[ht!]
    \centering
    \begin{subfigure}{.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{net-negative-0.1/full/figures/90_nodes/balance_map_carbon.png}
        \label{fig:capacity_map_carbon_full_nn}
    \end{subfigure}%
    \begin{subfigure}{.5\textwidth}
        \centering
        \includegraphics[width=\linewidth]{net-negative-0.1/full/figures/90_nodes/balance_map_hydrogen.png}
        \label{fig:capacity_map_hydrogen_full_nn}
    \end{subfigure}
    \caption{Optimal operation, flows and prices of the carbon (left) and hydrogen (right) sectors for the \hydrogenmodel{} in the net negative scenario. For each region, upper semicircles show the average production per technology, lower semicircles the consumption, and colors the average marginal prices. Carbon Sequestration offshore are drawn as full circles. Lines and arrows show the interregional transportation. With the tightened emission target, the \hybridmodel{} decreases the \hydrogengrid{} layout in comparison to the Net-Zero scenario and increases \carbon{} transport from inland point sources to the sequestration sites.
    }
    \label{fig:capacity_maps_full_nn}
\end{figure*}
